Highly Available Transactions: Virtues and Limitations

### 2013-03-30 writeup/evaluation.tex

With a firm understanding of which of several ACID and distributed
consistency properties are available and which are not, we begin to
understand their implications for existing algorithms, applications,
and deployments. Specifically:

@@

\vspace{.5em}\noindent\textbf{Serializablity} To establish a serial
order on transactions, all existing algorithms for achieving
serializability of general-purpose read-write transactions in a
distributed setting~\cite{bernstein-book} may require at least one
round trip time (RTT) before committing. As an example, traditional
two-phase locking for a transaction of length $T$ requires $T$
\texttt{lock} operations and at least one \texttt{unlock} operation.
Each of these operations requires coordination, either with other
database replicas or with a lock service. If the coordination
mechanism for locking is unavailable, transactions cannot
commit. Similarly, optimistic concurrency control requires contacting
a validation service, while deterministic transaction
scheduling~\cite{deterministic-scheduling} requires a
scheduler. Serializability under multi-version concurrency control
fares somewhat better but still requires checking for update
conflicts.

Perhaps unsurprisingly, the reliance on a globally agreed total order
necessitates a minimum of one round-trip to a designated master or
coordination service for each of these classic algorithms.  This is
intuitively a consequence of serializability's lack of high
availability~\cite{davidson-survey} (or, as we have shown, reliance on
preventing Lost Update and Write Skew).  The cost of the (minimum one)
round trip will be determined by the deployment environment, as we saw
in Section~\ref{sec:motivation}.

\vspace{.5em}\noindent\textbf{Non-serializablity} Perhaps
interestingly, many existing distributed implementations of weak
isolation are not highly available. Lock-based proposals such as those
used to provide weak isolation in Gray's original
proposal~\cite{gray-isolation} do not degrade gracefully in the
presence of partial failure. (Note, however, that lock-based protocols
\textit{do} offer the benefit of recency guarantees.) While
multi-versioned storage systems allow for a variety of transactional
guarantees, we have not seen proposals for truly weak isolation (e.g.,
non-``tentative update'' schemes) in this context. Chan and Gray's
read-only transactions provide read-only transactions with item-cut
isolation, causal consistency, and transactional atomicity but are
unavailable in the presence of coordinator failure~\cite{readonly},
similar to read-only and write-only transactions more recently
proposed by Eiger~\cite{eiger}.  Bolt-on causal
consistency~\cite{bolton}, Brantner's S3 database~\cite{kraska-s3},
and Bayou~\cite{sessionguarantees} can all provide variants of the
above model with high availability, but none provide this HAT
functionality without substantial modification; Swift~\cite{swift} is
closest to providing full HAT operation in a sticky
context. Similarly, the MDCC~\cite{mdcc} protocol offers Read
Committed isolation with Lost Update avoidance but is similarly
unavailable due to its reliance on preventing write conflicts.

As we have seen, it is possible to implement many guarantees weaker
than serializability---including guarantees that are achievable with
high availability---and not achieve high availability. We expect that
further understanding the trade-offs between semantic strength,
performance, and storage and metadata requirements will remain an area
of active research for the foreseeable future.

@@

Thus far, we have largely ignored the question of which applications
require semantics which are unavailable to HATs. As we showed in
Section~\ref{sec:hats}, the main cost of choosing HATs comes in the
inability to prevent Lost Update, Write Skew, and provide recency
bounds. In this section, we attempt to understand when these
guarantees matter both abstractly and by the use of a representative
customer order transaction application patterned off of the TPC-C
benchmark~\cite{tpcc}.

\vspace{.5em}\noindent\textbf{Commutativity and Monotonicity} Recent
work on the CALM Theorem~\cite{calm}, Lattice-based state
update~\cite{blooml}, and Commutative and Replicated Data
Types~\cite{crdt} demonstrates that, if updates logically commute,
then they can often be safely performed in different orders at
different locations in a distributed system. Accordingly, as long as
all writes are delivered to all replicas, then a system executing
monotonic logic with commutative operators may not suffer from
application-level consistency anomalies as a result of Lost Update or
Write Skew anomalies. However, applications with non-monotonic state
mutation will not, in general, be able to maintain application-level
consistency constraints. Similarly, applications requiring write
visibility between clients should opt for unavailability instead of
possibly missing a write visibility deadline.

\vspace{.5em}\noindent\textbf{TPC-C} Absent a necessary and sufficient
result describing exactly which application logics can be achieved
with high availability and to make these requirements more concrete,
we consider an example application: the TPC-C benchmark. Our goal is
\textit{not} to provide a performance analysis, as a HAT
configuration will be non-compliant. Rather, we examine the five
queries contained in the benchmarks---ostensibly representative of a
standard transaction processing application---to understand which are
achievable in a HAT context. Surprisingly, four of five can be
executed with HATs, while the fifth potentially requires unavailablity.

TPC-C consists of five transactions, capturing operation of a
real-world warehouse, including sales, payments, and deliveries. Two
transactions---\textit{Order-Status} and \textit{Stock-Level}---are
read-only and can be executed safely under HATs. Clients may read
stale data, but this does not violate TPC-C requirements and clients
will read their writes if they are sticky-available. Another
transaction type, \textit{Payment}, updates running balances for
warehouses, districts, and customer records as well as providing an
audit trail. The transaction is increment-only, so all balance
increase operations can commute, and TA allows the maintainence of any
foreign-key constraints (e.g., \texttt{UPDATE/DELETE CASCADE}).

@@

There are three challenges that only allow HATs to place non-compliant
New-Orders and to fail on Payments. First, New-Order requires a unique
ID number for each order. We can create a unique number by, say,
concatenating the client ID and a timestamp. However, the TPC-C
specification requires order numbers to be \textit{sequentially}
assigned, which requires Lost Update. Accordingly, HATs cannot provide
compliant TPC-C execution but can maintain uniqueness
constraints. Second, the New-Order transaction decrements inventory
counts: what if the count becomes negative? Fortunately, TPC-C
New-Order restocks each item's inventory count (increments by 91) if
it would become negative as the result of placing an order. This means
that, even in the presence of concurrent New-Orders, an item's stock
will never fall below zero. This is TPC-C compliant, but a HAT system
might end up with substantially more stock than in a single-site
implementation. Third, delivery events are non-monotonic---deleting a
pending New-Order transaction---and should be idempotent in order to
avoid billing a customer twice; this implies a need to prevent Lost
Update. We can avoid this issue by moving the non-monotonicity to the
real world---the carrier that picks up the package for an order can
ensure that she is the only carrier who has done so---but cannot
provide a correct execution with HATs alone.

>>>>

We now have an understanding of which guarantees are and are not
HAT-compliant. In this section, we begin to understand the impact of
this high availability on existing systems and applications as well as
briefly demonstrate HAT algorithms on public cloud
infrastructure. Specifically:

@@

\vspace{.5em}\noindent\textbf{Serializablity} To establish a serial
order on transactions, algorithms for achieving serializability of
general-purpose read-write transactions in a distributed
setting~\cite{bernstein-book} may require at least one round trip time
(RTT) before committing. As an example, traditional two-phase locking
for a transaction of length $T$ requires $T$ \texttt{lock} operations
and at least one \texttt{unlock} operation.  In a distributed
environment, each of these operations requires coordination, either
with other database replicas or with a lock service. If the
coordination mechanism for locking is unavailable, transactions cannot
commit. Similarly, optimistic concurrency control requires
coordinating via a validation step, while deterministic transaction
scheduling~\cite{deterministic-scheduling} requires a
scheduler. Serializability under multi-version concurrency control
also requires checking for update conflicts.

In summary, perhaps unsurprisingly, the reliance on a globally agreed
total order necessitates a minimum of one round-trip to a designated
master or coordination service for each of these classic algorithms.
This is intuitively a consequence of serializability's lack of high
availability~\cite{davidson-survey} (or, as we have shown, reliance on
preventing Lost Update and Write Skew).  The cost of the (minimum one)
round trip will be determined by the deployment environment, as we saw
in Section~\ref{sec:motivation}.

\vspace{.5em}\noindent\textbf{Non-serializablity} Many existing
distributed implementations of weak isolation are not highly
available. Lock-based proposals such as those used to provide weak
isolation in Gray's original proposal~\cite{gray-isolation} do not
degrade gracefully in the presence of partial failure. (Note, however,
that lock-based protocols \textit{do} offer the benefit of recency
guarantees.) While multi-versioned storage systems allow for a variety
of transactional guarantees, we have not seen proposals for truly weak
isolation (e.g., non-``tentative update'' schemes) in this
context. Chan and Gray's read-only transactions provide read-only
transactions with item-cut isolation, causal consistency, and
transactional atomicity (session PL-2L) but are unavailable in the
presence of coordinator failure~\cite{readonly}, similar to read-only
and write-only transactions more recently proposed by
Eiger~\cite{eiger}.  Bolt-on causal consistency~\cite{bolton},
Brantner's S3 database~\cite{kraska-s3}, and
Bayou~\cite{sessionguarantees} can all provide variants of session
PL-2L with high availability, but none provide this HAT functionality
without substantial modification; Swift~\cite{swift} is closest to
providing full HAT operation in a sticky context. Similarly, the
MDCC~\cite{mdcc} protocol offers Read Committed isolation with Lost
Update avoidance but is similarly unavailable due to its reliance on
preventing write conflicts.

As we have seen, it is possible to implement many guarantees weaker
than serializability---including guarantees that are achievable with
high availability---and still not achieve high availability.

@@

Thus far, we have largely ignored the question of which applications
require semantics that are unavailable to HATs. As we showed in
Section~\ref{sec:hats}, the main cost of choosing HATs comes in the
inability to prevent Lost Update, Write Skew, and provide recency
bounds. In this section, we attempt to understand when these
guarantees matter both abstractly and by the use of a representative
customer order transaction application patterned off of the TPC-C
benchmark~\cite{tpcc}.

\vspace{.5em}\noindent\textbf{Commutativity and Monotonicity} Recent
work on the CALM Theorem~\cite{calm}, Lattice-based state
update~\cite{blooml}, and Commutative and Replicated Data
Types~\cite{crdt} demonstrates that, if updates logically commute,
then they can often be safely performed in different orders at
different locations in a distributed system. Accordingly, as long as
all writes are delivered to all replicas, then a system executing
monotonic logic with commutative operators may not suffer from
application-level consistency anomalies as a result of Lost Update or
Write Skew anomalies. However, applications with non-monotonic state
mutation will not, in general, be able to maintain application-level
consistency constraints with HATs alone. Similarly, applications
requiring write visibility between clients should opt for
unavailability instead of possibly missing a write visibility
deadline.

\vspace{.5em}\noindent\textbf{TPC-C} Absent a necessary and sufficient
result describing exactly which application logics can be achieved
with high availability and to make these requirements more concrete,
we consider an example application: the TPC-C benchmark. Our goal is
\textit{not} to provide a performance analysis, as a HAT
configuration will be non-compliant. Rather, we examine the five
queries contained in the benchmarks---ostensibly representative of a
standard transaction processing application---to understand which are
achievable in a HAT context. Surprisingly, four of five can be
executed with HATs, while the fifth may require unavailablity.

TPC-C consists of five transactions, capturing operation of a
wholesale warehouse, including sales, payments, and deliveries. Two
transactions---\textit{Order-Status} and \textit{Stock-Level}---are
read-only and can be executed safely under HATs. Clients may read
stale data, but this does not violate TPC-C requirements and clients
will read their writes if they are sticky-available. Another
transaction type, \textit{Payment}, updates running balances for
warehouses, districts, and customer records as well as providing an
audit trail. The transaction is increment-only, so all balance
increase operations can commute, and TA allows the maintainence of any
foreign-key constraints (e.g., \texttt{UPDATE/DELETE CASCADE}).

@@

There are three challenges that only allow HATs to place non-compliant
New-Orders and to fail on Payments. First, New-Order requires a unique
ID number for each order. We can create a unique number by, say,
concatenating the client ID and a timestamp. However, the TPC-C
specification requires order numbers to be \textit{sequentially}
assigned, which requires preventing Lost Update. Accordingly, HATs
cannot provide compliant TPC-C execution but can maintain uniqueness
constraints. Second, the New-Order transaction decrements inventory
counts: what if the count becomes negative? Fortunately, TPC-C
New-Order restocks each item's inventory count (increments by 91) if
it would become negative as the result of placing an order. This means
that, even in the presence of concurrent New-Orders, an item's stock
will never fall below zero. This is TPC-C compliant, but a HAT system
might end up with substantially more stock than in a single-site
implementation. Third, delivery events are non-monotonic---deleting a
pending New-Order transaction---and should be idempotent in order to
avoid billing a customer twice; this implies a need to prevent Lost
Update. We can avoid this issue by moving the non-monotonicity to the
real world---the carrier that picks up the package for an order can
ensure that she is the only carrier who has done so---but cannot
provide a correct execution with HATs alone. However, according to
distributed transaction architects~\cite{helland}, these compensatory
actions are relatively common in real-world business processes.

### 2013-03-30 writeup/futurework.tex

\textbf{Where do semantics-based concurrency control and weakened
  failure models lie in the HAT landscape?} The vast amount of
literature on semantics-based concurrency control provides hints as to
when coordination can be avoided while providing ``strongly
consistent'' outcomes for limited failure scenarios. We have performed
a narrow case study in this work, but a general taxonomization of
techniques such as escrow transactions~\cite{escrow} and
Sagas~\cite{sagas} could help answer the challenge of programming in a
weakly consistent environment. Additionally, HAT guarantees can be
strengthened by considering failure models with, say, bounded
asynchrony or service-level agreements from underlying
hardware. Additionally, while we have focused on systems that provide
a single availability mode, there are several interesting hybrids to
consider.

>>>>

\textbf{Where do semantics-based concurrency control and weakened
  failure models lie in the HAT landscape?} The vast amount of
literature on semantics-based concurrency control provides hints as to
when coordination can be avoided while providing ``strongly
consistent'' outcomes for limited failure scenarios. We have performed
a narrow case study in this work, but a general taxonomization of
techniques such as escrow transactions~\cite{escrow} and
Sagas~\cite{sagas} could help answer the challenge of programming in a
weakly consistent environment. Additionally, HAT guarantees can be
strengthened by considering failure models with, say, bounded
asynchrony or service-level agreements from underlying
hardware. Additionally, while we have focused on systems that provide
a single availability mode, there are several interesting hybrids to
consider. For example, a system may provide ``best-effort'' stickiness
(or serializability!) in the absence of partitions, but, in the
presence of partitions, ``lose'' guarantees by having clients
reconnect to an available replica. This strategy provides stronger
guarantees in the absence of partitions but, to provide availability,
sacrifices them in the event of partitions.

### 2013-03-30 writeup/motivation.tex

Peter Deutsch begins his classic list of ``Fallacies of Distributed
Computing'' with two concerns fundamental to distributed database
systems: ``\textit{1.)}  The network is reliable. \textit{2.)} Latency
is zero''~\cite{fallacies-deutsch}. On a single server, communication
channels are robust and relatively fast, and few single-server systems
need to account for a possibility of partial system failure. However,
in a distributed setting, both of these assumptions are no longer
valid. To extend Jim Gray's storage latency analogy, if memory is 1.5
hours from the CPU and disk is 2 years away, then, especially across
wide-area networks (WANs), remote server accesses can exceed 30 years,
with the added concern that messages can be lost in transit or delayed
indefinitely by network partitions~\cite{gray-rules}. Accordingly,
ignoring network failures and latency can ``cause big trouble and
painful learning experiences''~\cite{fallacies-deutsch}.

In this section, we quantify the effect of network partitions and
latency in several real-world deployments. While the importance of
network behavior is a topic of considerable speculation within the
database community~\cite{stonebraker2010errors}, there is mounting
evidence---both anecdotal and experimental---that partitions occur and
network latencies are non-negligible. We subsequently discuss the
impact of these findings on modern database operation and motivate the
case for HATs.

@@

Network partitions---or, informally, communication link failures
between participants in a distributed system---are challenging. A
system facing network partitions faces the impossibility of
simultaneously maintaining a correct ``global'' view of system state
while simultaneously providing online operation to all of its
clients~\cite{davidson-survey}. For networks that do not partition,
maintaining semantic guarantees and available operation is
straightforward~\cite{stonebraker2010errors}. However, given the
\textit{possibility} of partitions, a system must be prepared to
either lose availability of operations or its ability to reason about
global state. In this section, we present evidence for the frequency
and non-triviality of WAN and LAN partitions.

@@

Fundamentally, the speed at which two servers can communicate is
(according to modern physics) bounded by the speed of light. Two
servers located on opposite sides of the Earth face, in the best
case---communicating via a hypothetical link crossing through the
center of the planet---require a minimum 85.1 ms RTT (133.7 ms if sent
at ground level). As services are replicated to multiple,
geographically distinct sites, the cost of communication increases to
this upper bound.

In real deployments, latencies are actually higher than the speed of
light dictates due to routing, congestion, and processing times; if
not, servers a kilometer apart would enjoy a 6.7 $\mu$s RTT. To
illustrate the difference between intra-datacenter, inter-datacenter,
and inter-planetary networks, we performed a measurement study of
network behavior on Amazon's EC2, a widely-used public compute
cloud. We measured one week of ping times between all seven EC2
geographic ``regions'', across three ``availability zones'' (closely
co-located datacenters), and within a single ``availabilty zone''
(datacenter), at a granularity of 1s (dataset to be released).

We summarize the results of our network measurement study in
Table~\ref{table:rtt}. On average, inter-datacenter communication (1c)
is between 1.82 and 6.38 times faster than across geographically
co-located datacenters (1b) and between 40 and 647 times faster than
across geographically distributed datacenters (1a). The cost of
wide-area communication exceeds the speed of light: for example,
communicating from S\~{a}o Paulo to Singapore is lower-bounded by
speed-of-light transit RTT of 106.7ms but, on average, incurs a
362.8ms RTT (95th percentile: 649ms). As shown in
Figure~\ref{fig:rtt}, the distribution of latencies varies between
links, but the trend---in line with other recent studies~\cite{mdcc,
  redblue}---is clear: remote communication has a substantial cost.

@@

Mitigating the effects of partitions and network latency is a
difficult task, requiring systems designers to choose between
fundamental trade-offs. In this section, we briefly discuss known
semantic limits for systems wishing to maintain high availability and
their impact on modern database systems. While the gold standard of
ACID semantics is unavailable, we show that many database systems
operate under much weaker models, motivating further study of these
models in a highly available context in the remainder of this paper.

One of the most popularly celebrated impossibility results in
distributed systems is Brewer's CAP Theorem, which, as formally proven
by Gilbert and Lynch, states that it is impossible to maintain
linearizability , or the ability to read the last completed write,
(``C,'' or ``consistency''---albeit poorly named) and replica
availability, or a guaranteed response, (``A'') in the presence of
network partitions (``P'')~\cite{gilbert-cap}. Brewer's Theorem has
roots in decades of distributed database
designs~\cite{davidson-survey} and has had a substantial impact on
recent large-scale distributed systems design: while CAP is narrowly
scoped to the impossibility of building highly available linearizable
systems, it is often interpreted to preclude a larger set of
guarantees, including ACID (transactional atomicity, consistency,
isolation and durability guarantees), as evidenced by the lack of
highly available systems providing them.

Database researchers and designers have long realized that
serializability---the gold standard of ACID isolation, which
guarantees application-level consistency---is not achievable in a
highly available system~\cite{davidson-survey}.  However, database
systems offer a range of ACID properties besides
serializability. Within a single-node database, the coordination
penalties associated with ensuring equivalence with a serial execution
can be severe and manifest themselves in the form of decreased
concurrency (and, subsequently, performance degradation, scalability
limitations, and, often, external aborts like
deadlocks)~\cite{gray-isolation}. Instead, databases offer a host of
so-called \textit{weak isolation} models that allow varying
restrictions on the space of schedules that are allowable by the
system~\cite{adya}. None of these weak isolation models guarantees
serializability, but the benefits of these models can often outweigh
costs of possible consistency anomalies that might arise from their
use.

To understand the prevalence of these weak isolation models, we
recently surveyed the default and maximum isolation models provided by
18 relational databases, often claiming to provide ``ACID'' or
``NewSQL'' functionality~\cite{hat-hotos}. As shown in
Table~\ref{table:existing}, only three out of 18 databases provided
serializability by default, and at least eight did not provide
serializability as an option at all. This is particularly surprising
when we consider the widespread deployment of many of these
non-serializable databases, like Oracle 11g, which are known to power
major businesses and product functionality. Given that these
transactional models are frequently used, our inability to provide
serializability in arbitrary HATs appears non-fatal for practical
applications. If application writers and database vendors have already
decided that the benefits of weak isolation outweigh potential
application inconsistencies, then, in a highly available environment
that prohibits serializability, similar decisions may be tenable.

@@

The primary challenge in providing weak isolation in a HAT context is
that, unfortunately, it is unclear \textit{which} of these ACID
(particularly isolation) guarantees can be provided with high
availability. The existing algorithms for providing weak isolation are
typically designed for a single-node context and are, to the best of
our knowledge, unavailable due to a reliance on concurrency control
mechanisms like locking that are not resilient to partial failure
(Section~\ref{sec:evaluation}). Morever, we are not aware of any prior
literature that provides guidance as to the relationship between weak
isolation and high availability: prior work has examined the
relationship between serializability and high
availability~\cite{davidson-survey} and weak isolation on a
single-server~\cite{adya, ansicritique} but never weak isolation and
high availability taken together. We believe that this apparent gap in
the database literature is a possible contributing factor to the
often-apparent attitude~\cite{hn, foundation-article} that
general-purpose transactional semantics are either too expensive to
provide in many modern distributed databases---except under special
circumstances, such as providing operations over groups of co-located
data items~\cite{entitygroup}.

>>>>

Distributed databases face a host of challenges that their
single-server counterparts are exempted from, largely stemming from
network asynchrony and partial failure. These challenges mean that
maintaining semantic guarantees has a cost, the exact price of which
is a subject of considerable speculation in the database
community~\cite{stonebraker2010errors}. In this section, we attempt to
quantify the prevalence and cost of network partitions and latency in
several real-world deployments. This data and a recent survey of
existing database semantics motivates the study of highly available
transaction systems.

(TODO-PDB: keep this?) We will begin by studying the behavior of
network partitions and latency. Peter Deutsch begins his classic list
of ``Fallacies of Distributed Computing'' with two concerns
fundamental to distributed database systems: ``\textit{1.)}  The
network is reliable. \textit{2.)} Latency is
zero''~\cite{fallacies-deutsch}. On a single server, communication
channels are robust and relatively fast, and few single-server systems
need to account for a possibility of partial system failure. However,
in a distributed setting, both of these assumptions are no longer
valid. To extend Jim Gray's storage latency analogy~\cite{gray-rules},
if memory is 1.5 hours from the CPU and disk is 2 years away, then,
especially across wide-area networks (WANs), remote server accesses
can exceed 30 years, with the added concern that messages can be lost
in transit or delayed indefinitely by network partitions. Accordingly,
ignoring network failures and latency can ``cause big trouble and
painful learning experiences''~\cite{fallacies-deutsch}.

@@

Network partitions---informally, communication failures between
participants in a distributed system---are challenging. A system
facing network partitions faces the impossibility of simultaneously
maintaining a correct ``global'' view of system state while
simultaneously providing online operation to all of its
clients~\cite{davidson-survey}. For networks that do not partition,
maintaining semantic guarantees and providing ``always on'' operation
is straightforward~\cite{stonebraker2010errors}. However, given the
\textit{possibility} of partitions, a system must be prepared to
either lose availability of operations or its ability to reason about
global state. In this section, we present evidence for the frequency
and non-triviality of WAN and LAN partitions.

@@

Fundamentally, the speed at which two servers can communicate is
(according to modern physics) bounded by the speed of light. Two
servers located on opposite sides of the Earth face, in the best
case---communicating via a hypothetical link crossing through the
center of the planet---require a minimum 85.1 ms RTT (133.7 ms if sent
at ground level). As services are replicated to multiple,
geographically distinct sites, the cost of communication increases to
this lower bound. In real deployments, transit times are actually
slower than the speed of light due to routing, congestion, and
processing times. To illustrate the difference between
intra-datacenter, inter-datacenter, and inter-planetary networks, we
performed a measurement study of network behavior on Amazon's EC2, a
widely-used public compute cloud. We measured one week of ping times
between all seven EC2 geographic ``regions'', across three
``availability zones'' (closely co-located datacenters), and within a
single ``availabilty zone'' (datacenter), at a granularity of 1s
(dataset to be released).

We summarize the results of our network measurement study in
Table~\ref{table:rtt}. On average, intra-datacenter communication (1c)
is between 1.82 and 6.38 times faster than across geographically
co-located datacenters (1b) and between 40 and 647 times faster than
across geographically distributed datacenters (1a). The cost of
wide-area communication exceeds the speed of light: for example,
communicating from S\~{a}o Paulo to Singapore is lower-bounded by
speed-of-light transit RTT of 106.7ms but, on average, incurs a
362.8ms RTT (95th percentile: 649ms). As shown in
Figure~\ref{fig:rtt}, the distribution of latencies varies between
links, but the trend---in line with other recent studies~\cite{mdcc,
  redblue}---is clear: remote communication has a substantial cost.

@@

Systems designers must make fundamental trade-offs between latency,
system behavior during partitions, and achievable semantics. In this
section, we briefly discuss known semantic limits for systems wishing
to maintain high availability and their impact on modern database
systems. We show that many database systems operate under weak model
whose availability characteristics are unknown, motivating further
study of these models in a highly available context in the remainder
of this paper.

One of the most popularly celebrated impossibility results in
distributed systems is Brewer's CAP Theorem, which, as formally proven
by Gilbert and Lynch, states that it is impossible to maintain
linearizability , or the ability to read the last completed write,
(``C,'' or ``consistency''---albeit poorly named) and replica
availability, or a guaranteed response, (``A'') in the presence of
network partitions (``P'')~\cite{gilbert-cap}. Brewer's Theorem has
roots in decades of distributed database
research~\cite{davidson-survey} and has had a substantial impact on
recent large-scale distributed systems design: while CAP is narrowly
scoped to the impossibility of building highly available linearizable
systems, it is often interpreted to preclude a larger set of
guarantees, including ACID (transactional atomicity, consistency,
isolation and durability guarantees)~\cite{hn, foundation-article} as
also evidenced by the lack of highly available systems providing them.

Database researchers and designers have long realized that
serializability---the gold standard of ACID isolation, which
guarantees application-level consistency---is not achievable in a
highly available system~\cite{davidson-survey}.  However, to increase
concurrency, database systems offer a range of ACID properties weaker
than serializability. Within a single-node database, the coordination
penalties associated with serializability can be severe and manifest
themselves in the form of decreased concurrency (and, subsequently,
performance degradation, scalability limitations, and, often, external
aborts like deadlocks)~\cite{gray-isolation}. Instead, databases offer
a host of so-called \textit{weak isolation} models that allow varying
restrictions on the space of schedules that are allowable by the
system~\cite{adya}. None of these weak isolation models guarantees
serializability, but, as we see below, the benefits of these models
are often considered to outweigh costs of possible consistency
anomalies that might arise from their use.

To understand the prevalence of these weak isolation models, we
recently surveyed the default and maximum isolation models provided by
18 databases, often claiming to provide ``ACID'' or
``NewSQL'' functionality~\cite{hat-hotos}. As shown in
Table~\ref{table:existing}, only three out of 18 databases provided
serializability by default, and at least eight did not provide
serializability as an option at all. This is particularly surprising
when we consider the widespread deployment of many of these
non-serializable databases, like Oracle 11g, which are known to power
major businesses and product functionality. Given that these
transactional models are frequently used, our inability to provide
serializability in arbitrary HATs appears non-fatal for practical
applications. If application writers and database vendors have already
decided that the benefits of weak isolation outweigh potential
application inconsistencies, then, in a highly available environment
that prohibits serializability, similar decisions may be tenable.

@@

It is unknown \textit{which} of these ACID (particularly isolation)
guarantees can be provided with high availability. The existing
algorithms for providing weak isolation are typically designed for a
single-node context and are, to the best of our knowledge, unavailable
due to a reliance on concurrency control mechanisms like locking that
are not resilient to partial failure
(Section~\ref{sec:evaluation}). Morever, we are not aware of any prior
literature that provides guidance as to the relationship between weak
isolation and high availability: prior work has examined the
relationship between serializability and high
availability~\cite{davidson-survey} and weak isolation on a
single-server~\cite{adya, ansicritique} but never weak isolation and
high availability taken together.
